Formalizing Commutative Algebra in Coq: \\ Nakayama's Lemma\thanks{Source code for this work is available on the following site: \url{https://github.com/ku-sldg/algebra}}

### 2022-11-04 nakayama/paper.tex
Added figure, abstract, and made some prose edits

\title{ 
Formal Commutative Algebra in Coq: \\ Nakayama's Lemma\thanks{Source code for this work is available on the following site: \url{https://github.com/ku-sldg/algebra}}
}

@@

%\author{}
\author{ 
	Authors  
%	
	\vspace{.2cm}		
	\\ 
%	
	Institute for Information Sciences \\
	The University of Kansas \\
	Lawrence, KS 66045
	}
%\email{}

@@

\begin{abstract}
\emily[inline]{Abstract here }
\end{abstract}

@@

Our goal is to newly formalize theoretical, rather than computational, commutative algebra in Coq. 
We formally prove \emph{Nakayama's lemma} \cite{nakayama-1951, azumaya}, an essential result in the field. 
In doing so, we formalize algebraic structures that are fundamental to higher-level algebra, such as 
\emph{local rings} and \emph{(finitely generated) modules over commutative rings}. 
Rather than build upon some of the basic objects from abstract algebra, such as groups and rings, that have been formalized in Coq, e.g., in the  
\emph{Mathematical Components Library} 
\cite{mathcomp}, we start from scratch.

@@

@@

\begin{enumerate}[leftmargin=2cm,topsep=0cm,itemsep=0cm]
\item  $r \cdot (u + v)  = r \cdot u + r \cdot v$
\item $(r+s) \cdot u = r \cdot u + s \cdot u$
\item $(rs) \cdot u = r \cdot (s \cdot u)$
\item $1 \cdot u = u$ 
\end{enumerate}

@@

In order to state Nakayama's Lemma, we explain some notation:  If $I$ is an ideal of a commutative ring $R$ and 
 $M$ is an $R$-module, then $IM$ is the set of elements of the form 
	$a_1 u_1 + a_2 u_2 + \cdots + a_k u_k$, where, for some positive integer $k$, $a_1, \ldots, a_k \in I$ and $u_1, \ldots, u_k \in M$.  
Notice that due to the absorption property of ideals, $IM$ is an $R$-module contained in $M$.

If an $R$-module $M$ consists of only one element, this element must be its additive identity $0$ as an abelian group under addition.  We use the notation $M=0$ in this case.

@@

There are several alternate statements of Nakayama's Lemma that remove the hypothesis that $R$ must be local.  One can replace the unique maximal ideal with the Jacobson radical of the ring, which is the intersection of all maximal ideals. Alternatively, $I$ is an arbitrary proper ideal of a commutative ring $R$ and  
 $M$ is a finitely generated $R$-module for which $M=I M$, then this ensures the existence of a ring element $r$ congruent to $1$ modulo $I$ such that $rM = 0$, i.e., $r u = 0$ for every $u \in M$.

@@

@@

@@

\section{Formal Proof of Nakayama's Lemma}

>>>>

\title{ 
Formalizing Commutative Algebra in Coq: \\ Nakayama's Lemma\thanks{Source code for this work is available on the following site: \url{https://github.com/ku-sldg/algebra}}
}

@@

%\author{}
\author{ 
	Andrew Cousino \\ {\tt acousino@ku.edu}
	\and 
	Emily E.\ Witt \\ {\tt witt@ku.edu}	
	\and 
	Perry Alexander \\ {\tt palexand@ku.edu}
}	
	%

\vspace{-.3cm}

{\large
\begin{center}
	Institute for Information Sciences \\
	The University of Kansas \\
	Lawrence, KS 66045
\end{center}
}

@@

\todo[inline]{To do:  Add to/alter author list, title, and/or abstract as necessary}

\begin{abstract}
We describe our formal proof of Nakayama's Lemma, a fundamental theorem in the mathematical field of commutative algebra. The statement and proof of this result involve several commutative-algebraic structures including commutative rings, ideals of these rings, and modules over them, and we also explain our process of formalizing these structures. 
\end{abstract}

@@

Our goal is to newly formalize theoretical, rather than computational, commutative algebra in Coq. 
We formally prove \emph{Nakayama's lemma} \cite{nakayama-1951, azumaya}, an essential result in the field. 
In doing so, we formalize algebraic structures that are fundamental to higher-level algebra, such as 
\emph{local rings} and \emph{(finitely generated) modules over commutative rings}. 
Rather than build upon some of the basic objects from abstract algebra, such as groups and rings, that have been formalized in Coq, e.g., in the  
\emph{Mathematical Components Library} 
\cite{mathcomp}, we start from scratch. \todo{To do: Verify whether MathComp only formalized finite algebraic structures}
 The theory, including the formalization of all algebraic structures, is approximately  $100$ kB and \textcolor{red}{??}\todo{To do:  Add number} lines of code.

@@

\emily[inline]{Emily: It might be more consistent to require a ring to contain $1$ in general, instead of adding this to the axioms of a commutative ring.  Perhaps we could do this in the code, and then here, without too much effort? }

@@

\begin{enumerate}[leftmargin=6cm,topsep=0cm,itemsep=0cm]
\item  $r \cdot (u + v)  = r \cdot u + r \cdot v$
\item $(r+s) \cdot u = r \cdot u + s \cdot u$
\item $(rs) \cdot u = r \cdot (s \cdot u)$
\item $1 \cdot u = u$ 
\end{enumerate}

@@

In order to state Nakayama's Lemma, we first explain some notation:  If $I$ is an ideal of a commutative ring $R$ and 
 $M$ is an $R$-module, then $IM$ is the set of elements of the form 
	$a_1 u_1 + a_2 u_2 + \cdots + a_k u_k$, where, for some positive integer $k$, $a_1, \ldots, a_k \in I$ and $u_1, \ldots, u_k \in M$.  
Notice that due to the absorption property of ideals, $IM$ is an $R$-module contained in $M$.

If an $R$-module $M$ consists of only one element, this element must be its additive identity $0$ as an abelian group under addition.  The notation $M=0$ means that we are in this situtation.

@@

There are alternate statements of Nakayama's Lemma that remove the hypothesis that $R$ must be local.  One can replace the unique maximal ideal with the Jacobson radical of the ring, which is the intersection of all maximal ideals. Alternatively, $I$ is an arbitrary proper ideal of a commutative ring $R$ and  
 $M$ is a finitely generated $R$-module for which $M=I M$, then this ensures the existence of a ring element $r$ congruent to $1$ modulo $I$ such that $rM = 0$, i.e., $r u = 0$ for every $u \in M$.

@@

We start by describing the process of formalization of algebraic structures, and with that in hand, then move on to the formal proof of Nakayama's lemma.

@@

\begin{figure}[t]
\caption{The hierarchy of our algebraic structures}
\includegraphics[width=.4\textwidth]{algebraic-structures.pdf}
\centering
\todo[inline]{To do:  Check source code to verify whether the hierarchy figure should be altered (subgroups/other subobjects, quotients, finitely generated modules, etc.)}
\end{figure}

@@

\subsection{Formal Proof of Nakayama's Lemma}
